The impacts of demography on linked selection in maize

### 2015-07-29 manuscript.tex
updates... "complete", depending on your def...

\subsection{Demography strongly impacts efficiency of purifying selection}
Explain that singleton patterns are flipped compared to pi. Put this in the context of large pop recent, small pop past, so efficient selection recent, inefficient selection past. Explain the importance that this has on old deleterious alleles (potential for many due to drift during bottleneck) and recent deleterious alleles (efficiently purged from population).

>>>>

\subsection{Demography strongly impacts efficiency of purifying selection}
The observation that maize pairwise diversity is less impacted by the distance to the nearest gene than is teosinte pairwise diversity is reasonable from a long-term evolutionary standpoint. Theory demonstrates that purifying selection is more efficient in a large population than in a small one \cite{somebody}, and this observation most likely reflects that prediction. More specifically, if teosinte $N_e$ remained relatively constant while maize bottlenecked and recovered exponentially, this provides that the average $N_e$ of maize over the previous several thousand generations is much smaller than that of teosinte, regardless of how much maize has ballooned in the recent past. Therefore, our observation shows that purifying selection in maize has not purged deleterious alleles, or the neutral alleles they are linked to, as effectively in maize as in teosinte.

The reversal of this trend when we analyze only singleton diversity instead of pairwise diversity, however, stands out as a notable observation. Every mutation begins as a singleton (an allele present in only one individual), and therefore singletons are, on average, the youngest class of alleles that we can observe. Therefore unlike pairwise patterns of diversity, which are most heavily influenced by intermediate frequency alleles based on the definition of $\pi$ \cite{nei1979}, singleton diversity is most influenced by recent patterns of evolution. Hence, because our demographic estimation indicates dramatic expansion of maize $N_e$ in the recent past, we expect for purifying selection to presently operate more efficiently in maize than in teosinte. This observation is very clearly demonstrated by the fact that singleton diversity in maize is more impacted by the distance to the nearest gene than it is in teosinte, as was shown in Figure \ref{fig:purify}.

A consequence of the inefficient purifying selection that maize experienced during its bottleneck is likely that it harbors more deleterious alleles segregating at intermediate frequency than does teosinte. This could be a part of the explanation of why maize inbreds have continued to improve over the past several decades \cite{something}; if deleterious alleles tend to be recessive and are particularly frequent, they will have increased opportunities to display their phenotypes in inbreds. These results also demonstrate that recent purifying selection in maize has become much more effective, potentially explaining the ongoing improvement of these inbreds as maize lines are continuously selected. The sample we studied contained only open-pollinated varieties of both maize and teosinte, so we cannot explicitly test this prediction.

Ultimately, we have shown that purifying selection in maize has operated very differently than purifying selection in teosinte. This observation along with the potential for soft sweeps appear to explain the phenotypic divergence between maize and teosinte much more completely than do hard sweeps on coding variation. Importantly, our estimation of the parameters of the maize domestication bottleneck contribute to the understanding of how the deomography of crop domestication can impact crop diversity. The bottleneck-effects from a sudden collapse in population size have been well studied and are known to impact crops for thousands of generations. Our observations demonstrate that the rapid expansion experienced by many crops after domestication can also have a profound influence on patterns of diversity, and the effects of this expansion should be accounted for as important contributors to long-term evolution.
